Flabby and injective objects in toposes

### 2022-02-14 paper-flabby-objects.tex
Rewritten conclusion

A further motivation for this note was our desire to seek a constructive
account of sheaf cohomology. Sheaf cohomology is commonly defined using
injective resolutions, which can fail to exist in the absence of the axiom of
choice~\cite{blass:inj-proj-axc}, but flabby resolutions
can also be used in their stead, making them the obvious candidate for a
constructively sensible replacement of the usual definition. However, we show
in Section~\ref{sect:in-eff} and in Section~\ref{sect:conclusion} that flabby
resolutions present their own challenges, and in summary we failed to reach this
goal. The problem of giving a constructive account
of sheaf cohomology is still open.

@@

Can this issue be remedied by employing flabby resolutions
instead of injective ones? Classically, it is known that they can be used in
their stead. For instance, George Kempf developed the foundations of the
cohomology of quasicoherent sheaves on this premise~\cite{kempf:cohomology}. There are, however, two problems with this
suggestion. Firstly, all proofs known to us that flabby sheaves are
acyclic for the global sections functor require Zorn's lemma. This problem might
be mitigated by relying on the substitute property discussed
following Scholium~\ref{scholium:exact-as-presheaves}. \XXX{XXX}

\begin{enumerate}
\item \XXX{Reformulate as remark somewhere}

\item There appears to be some tension regarding the \effective topos:
Proposition~\ref{prop:semienough-flabby-modules} shows that at
least~$\neg\neg$-separated modules in the \effective topos always embed into
flabby modules, assuming the law of excluded middle in the metatheory, while
Proposition~\ref{prop:flabby-effective-sets} shows that no nontrivial \effective
module is flabby.
\end{enumerate}

We currently believe that it is not possible to give a constructive account of
a global cohomology functor which would associate to any sheaf of modules its
cohomology. However, it should be possible to do so for a restricted class of
sheaves, while still preserving the good formal properties expected from
derived functors.

@@

\XXX{cite more: Hyland §5,
Fourman-Scott 1979,
Johnstone 1977}

\XXX{quasicompact case}

@@

XXX: injective as module implies functionally flabby

Cite! https://arxiv.org/pdf/2008.10677.pdf

@@

1. Zorn
2. effective
3. predicative, no AU's

>>>>

A further motivation for this note was our desire to seek a constructive
account of sheaf cohomology. Sheaf cohomology is commonly defined using
injective resolutions, which can fail to exist in the absence of the axiom of
choice~\cite{blass:inj-proj-axc}, but flabby resolutions
can also be used in their stead, making them the obvious candidate for a
constructively sensible replacement of the usual definition. However, we show
in Section~\ref{sect:in-eff} and in Section~\ref{sect:conclusion} that flabby
resolutions present their own challenges, and in summary we failed to reach this
goal. The problem of giving a constructive account
of sheaf cohomology is still open~\cite{tenorio-mariano:cohomology}.

@@

Can this issue be remedied by employing flabby resolutions
instead of injective ones? Classically, it is known that they can be used in
their stead, and moreover George Kempf developed the foundations of the
cohomology of quasicoherent sheaves on this premise~\cite{kempf:cohomology}. There are, however, three issues with this
suggestion.

Firstly, all proofs known to us that flabby sheaves are
acyclic for the global sections functor require Zorn's lemma. This problem might
be mitigated by relying on the substitute property discussed
following Scholium~\ref{scholium:exact-as-presheaves}, but still acyclitity in
the usual sense will most likely not be attainable.

Secondly, while we have shown in Theorem~\ref{thm:enough-flabby-modules} that
enough flabby envelopes exist constructively, the analysis in Section~\ref{sect:in-eff}
demonstrates that the presented envelopes are in a sense deeply uncomputable.
Intuitionistic Zermelo--Fraenkel set theory~\cite{crosilla:cst-izf} and the
type theory of toposes~\cite{maietti:modular} do verify that they exist, but
only by virtue of excessive reliance on powersets, precluding concrete
computations; the construction is highly \emph{impredicative}, not meeting the
bar of \emph{predicative mathematics}~\cite{crosilla:predicativity}.

As a result, thirdly, flabby resolutions do not generalize from toposes to the
setting of \emph{arithmetic universes}, the predicative cousins of toposes
introduced by André Joyal which have recently been an important object of
consideration by Milly Maietti and
Steve Vickers~\cite{maietti:au,maietti-vickers:induction,vickers:sketches}.
These are not only interesting on their own and as a convenient foundation
for ensuring predicativity (and hence computability in a strict sense), but also
because they can be used to obtain base-independent proofs for the topos
case~\cite{vickers:classifying,hazratpour:phd}: \emph{Computing cohomology of
concrete spaces should not require fixing a universe of sets first.}

We currently believe that it is not possible to give a constructive account of
a global cohomology functor which would associate to any sheaf of modules its
cohomology. However, it should be possible to do so for a restricted class of
sheaves, while still preserving the good formal properties expected from
derived functors. To this end, a mix of the approaches using
pointwise Kan extensions, as cogently argued for by Emily
Riehl~\cite[Chapter~2]{riehl:cathtpy}, and ind-objects as presented in the Stacks
Project~\cite[Tag~05S7]{stacks-project} seems promising.

This framework is sufficiently flexible to not demand derived functors to be
defined everywhere. Rather, they will be defined just for those objects for
which we happen to have a suitable resolution. Only classically, by using
injective or flabby sheaves, can we pretend that every object has such a
resolution. We hope to report on details in future work.

@@

@@

@@

* quasicompact/spectral case
